---
title: '"Acceptabiliy Analysis for Catalan and Spanish Ratings"'
author: "Anna"
date: "`r Sys.Date()`"
output: github_document
---

**About this document:**
To render, run `rmarkdown::render(input = "/home/ansost/repos/CaSpaSyn/code/accepatibility-analysis.rmd", output_file = "/home/ansost/repos/CaSpaSyn/data/myfile.html")` in an R environment. 
This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
install.packages('tidyverse')
install.packages("devtools")
library(readr)
```

# Data exploration 
```{r}
df.initial <- read_csv('data/catalan.csv')
str(df.initial)
```

##  Mean Ratings 

### Active Ratings 

### Pronominal Ratings 

### p-Wert
Zweiseitiger t-Test die erste Idee, um mit einem Alpha von 0,05 zu schauen, ob es signifikante Unterschiede in den Ratings gab.

Später vielleicht: ob jedes Verb für sich signifikante Unterschiede bei Active vs. Pronominal aufweist, aber das kann man ja noch schauen

## Including Plots
